Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-04-30 writeups/paper.tex
check in writeup w/ fft

When the fluid absorbs the incident wave, it absorbs the incident horizontal
momentum flux as well, which is converted into additional horizontal momentum of
the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
phase velocity of the incident wave, the critical layer must instead propagate
downwards to accommodate the incident momentum. The total horizontal momentum of
the shear flow then obeys conservation equation
\begin{equation}
    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}(z, t)\;\mathrm{d}z
        \Delta S = 0.
\end{equation}
We define $\Delta S$ to be the change in flux across the critical layer, or
equivalently the negative of the absorbed flux. Treating $\bar{U}(z > z_c) =
\bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
\begin{equation}
    \rho(z_c) \bar{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
\end{equation}
$\Delta S(z_c)$ denotes the change in $S$ across $z_c$. For constant $\Delta S$
in time and $\rho \approx \rho_0$, this has analytical solution
\begin{equation}
    z_c(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{-\Delta S}
        .\label{eq:zc_sol}
\end{equation}
Note that constant $-\Delta S$ corresponds physically to the picture where
constant $S$ carried by the IGWs, given by \autoref{eq:S_lin}, is fully absorbed
at the critical layer, in which case $\Delta S = -S$.

@@

@@

Finally,

\subsection{Applicability to Astrophysical Systems}

% magnetic winding, radiative diffusivity + effective Re. Might be applicable to
% inwards focusing too

\subsection{Energy Transfer}

>>>>

When the fluid absorbs the incident wave, it absorbs the incident horizontal
momentum flux as well, which is converted into additional horizontal momentum of
the shear flow. Since the shear flow cannot exceed $\bar{U}_c$ the horizontal
phase velocity of the incident wave, the critical layer must instead propagate
downwards to accommodate the incident momentum. The total horizontal momentum of
the shear flow then obeys conservation equation
\begin{equation}
    \pd{}{t}\int\limits_0^{L_z} \rho(z) \bar{U}(z, t)\;\mathrm{d}z
        + \Delta S = 0.
\end{equation}
We define $\Delta S$ to be the change in flux across the critical layer, or
equivalently the negative of the absorbed flux. Treating $\bar{U}(z > z_c) =
\bar{U}_c, \bar{U}(z < z_c) = 0$ gives us exactly
\begin{equation}
    \rho(z_c) \bar{U}_c\pd{z_c}{t} = \Delta S(z_c(t)).\label{eq:zc_anal}
\end{equation}
$\Delta S(z_c)$ denotes the change in $S$ across $z_c$. For constant $\Delta S$
in time and $\rho \approx \rho_0$, this has analytical solution
\begin{equation}
    z_c(t) = -H\ln t - H\ln \frac{H\rho_0(z = 0)c_{ph, x}}{-\Delta S}
        .\label{eq:zc_sol}
\end{equation}
Note that constant $-\Delta S$ corresponds physically to the picture where
constant $S$ carried by the IGWs, given by \autoref{eq:S_lin}, is fully absorbed
at the critical layer, in which case $\Delta S = -S$.

@@

Finally, we may be interested in the power spectrum of the transmitted
horizontal momentum flux. Define:
\begin{align}
    \tilde{S}_>(k_x, t) = \abs{
        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_>(x, z_c - \Delta z, t)\;\mathrm{d}x}
            \label{eq:sfft_above},\\
    \tilde{S}_<(k_x, t) = \abs{
        \int\limits_0^{L_x} e^{ik_xx}\hat{S}_<(x, z_c + \Delta z, t)\;\mathrm{d}x}.
            \label{eq:sfft_below}
\end{align}
These values are plotted in \autoref{fig:nl_fft}. Note that we expect
$\tilde{S}_<(0, t) = S_<(t) = S_i(t) + S_d(t)$ and the same for $\tilde{S}_>$.
This is rather accurate for $\tilde{S}_<$ but less so for $\tilde{S}_>$; this is
since $S_>(t)$ decays very rapidly above $z_c$. Note furthermore that at later
times, the transmitted flux increases, per expectation, and shifts towards
higher $k_x$. Repeating this plot at higher viscosities shows significantly
lower $\tilde{S}_>$, showing that the $k_x$ of the transmitted waves still
changes greatly at the highest $\mathrm{Re}$ we are able to probe. Finally, note
that the area under $\tilde{S}$ changes with time; this is okay since we are
measuring $S$ at constant position, and different $k_x$ components of $S$ will
translate at their separate group velocities.

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fft.png}
    \caption{The power spectrum of the horizontal momentum flux at fixed
    distances below and above critical layer $z_c$ at certain times $t$ for
    simulation \texttt{0p5\_shres}. \autoref{eq:sfft_above} and
    \autoref{eq:sfft_below} are depicted here, with $\tilde{S}_<$ in solid lines
    and $\tilde{S}_>$ in dashed lines. Note that there is more incident flux
    than transmitted, and that transmitted flux roughly increases over
    time.}\label{fig:nl_fft}
\end{figure}

@@

Finally, we consider the impact of magnetic winding. In \citep{bukart}, magnetic
winding is used to enforce solid body rotation on the grounds that $t_A \gg
t_{gw}$, where
\begin{equation}
    t_A = \int\limits_0^R \frac{\sqrt{4\pi \rho}}{B_0}\;\mathrm{d}r
        \sim 10^2\;\mathrm{yr}\p*{\frac{10^3\;\mathrm{G}}{B}}.
\end{equation}
the Alfv\'en wave crossing time (evaluated for a CO WD in \citep{fullerIV})
measures the magnetic coupling time and $t_{gw}$ measures the gravitational wave
inspiral timescale. Before solid body rotation is attained, another relevant
timescale is the synchronization timescale $t_s$. For a tidal torque $\tau$ and
tidal forcing frequency $\sigma = m\p*{\Omega - \Omega_{spin}}$, we note that
angular momentum transfer is $\pd{M_{sync}}{t} \sigma R^2 = \tau$, where
$M_{sync}$ denotes the mass of the WD that has synchronized. Thus, the
synchronization timescale is
\begin{align}
    t_{sync} &\sim \frac{M_{sync}\sigma R^2}{\tau},\nonumber\\
        &\sim 2 \times 10^5\;\mathrm{yr}
            \p{\frac{M}{M_{\odot}}}
            \p{\frac{\sigma}{2\pi / (1\;\mathrm{hr})}}
            \p{\frac{R}{R_{\oplus}}}^2
            \p{\frac{10^{-14} GM_{\odot}^2/R_{\oplus}}{\tau}}.
\end{align}
A representative $\tau$ has been taken from \citep{bukart}. However, since only
the outer $\sim 10^{-4}M_{\odot}$ need be heated for the thermodynamically
interesting effects studied in \citep{fullerIV} and \citep{tidal_novae}, it
seems that magnetic winding cannot absolutely rule out energetic outbursts such
as tidal novae resulting from strong shear flows.

\subsection{Applicability to Other Astrophysical Systems}

[TODO flesh out]

\begin{itemize}
    \item $k_x, k_z$: In astrophysical systems, $k_{\perp} \ll k_r$. While we do
        not explore different $k_{1x}, k_{1z}, \omega_1$ in this study, with
        outgoing boundary conditions, there appears at first to be no $z$ length
        scale other than $k_{1z}$, so our results would seem to be invariant
        under rescaling of the $z$ length scale. However, true turbulence is
        expected to be isotropic at small scales, which may couple $k_{1x},
        k_{1z}$ in a way that $k_{1x} \ll k_{1z}$ produces different dynamics
        than $k_{1x} \lesssim k_{1z}$ as we've studied. This is a numerically
        difficult regime though, so we defer consideration to future work.

    \item Validity of plane-parallel approximation? We're all at $\gtrsim
        0.9R_{WD}$.

    \item Solar-type stars (inner convective, outer radiative): different
        equation of state/stratification but could be qualitatively similar.

    \item Solar-type stars: In~\cite{barker_ogilvie}, inwards-propagating IGW
        are excited that break via geometric focusing and effect
        synchronization. They find no reflected wave despite their nonlinear
        timescales being $10\times$ shorter than their viscous timescale.

        It is not immediately clear whether our results here apply when the flux
        is geometrically focused, but as a hypothesis we assume the convergence
        in \autoref{fig:agg} applies under geometric focusing as a zeroth
        approximation, perhaps as a property of the fluid motion within the
        geometrically thin critical layer.

        Associating $t_{} \sim \nu k^2 \approx \nu k_z^2$ with the visocus
        timescale and $t_{NL} \sim \vec{u} \cdot \vec{\nabla} \sim \omega$ for
        the nonlinear timescale, we find their $\lambda = \frac{t_{NL}}{t_L}
        \sim \mathrm{Re}$ our Reynolds number. Our simulations indicate
        $\mathrm{Re} \gtrsim 500$ are required to observe the correct asymptotic
        behavior in terms of horizontal momentum flux reflection/transmission,
        so it is possible their lack of reflection is viscosity limited.
\end{itemize}

\subsection{Heating}
